---
title: "Analysis of In/Out Experiment Data from One Subject"
author: "Robert Sekuler"
date: "`r Sys.Date()`"
output: html_document
params:
  show_code: FALSE  # Set to TRUE to display code
---

```{r setup, include=FALSE}
# Load required packages
library(readr)
library(dplyr)
library(knitr)
library(openxlsx)  # for Excel output
library(stringr)

# Control whether code is shown
the_echo <- params$show_code
knitr::opts_chunk$set(echo = the_echo)

# Define path to your CSV file
data_file <- "~/Desktop/rs_inORoutExpt_2025-04-28_16h43.25.066.csv"
```

## 1. Read, select data, and extract metadata
- Reads in the CSV file, skipping empty rows, selects the relevant columns, and removes separator lines.
- Parses the filename to extract `subject_id`, `run_date`, and `run_time`.

```{r read-data}
# Read CSV, select first 8 columns, filter out empty lines
dat <- read_csv(
  data_file,
  skip_empty_rows = TRUE
) %>%
  select(
    blocks.thisN,
    trials.thisN,
    StimulusPresented,
    stimulusCategory,
    BlurLevel,
    choice.keys,
    correct,
    choice.rt
  ) %>%
  filter(!is.na(StimulusPresented))

# Extract metadata
fname      <- basename(data_file)
no_ext     <- tools::file_path_sans_ext(fname)
parts      <- str_split(no_ext, "_")[[1]]
subject_id <- parts[1]
run_date   <- parts[2]
run_time   <- parts[3]

# Display metadata
cat("**Subject:**", subject_id, "  ",
    "**Run date:**", run_date, "  ",
    "**Run time:**", run_time, "\n")
```

## 2. Define conditions
- Assigns each trial to one of four conditions: indoor_sharp, indoor_blurred, outdoor_sharp, outdoor_blurred.

```{r define-conditions}
dat <- dat %>%
  mutate(
    condition = case_when(
      stimulusCategory == "i" & BlurLevel == 0 ~ "indoor_sharp",
      stimulusCategory == "i" & BlurLevel == 1 ~ "indoor_blurred",
      stimulusCategory == "o" & BlurLevel == 0 ~ "outdoor_sharp",
      stimulusCategory == "o" & BlurLevel == 1 ~ "outdoor_blurred",
      TRUE ~ NA_character_
    )
  )
```

## 3. Summarize across all blocks by condition
- Computes trial count, percent correct, and RT statistics (mean, median, SD) using only correct trials.
- Rounds all numeric results to four decimal places and adds `subject` column.

```{r summarize-condition}
summary_by_cond <- dat %>%
  group_by(condition) %>%
  summarize(
    n_trials     = n(),
    prop_correct = round(mean(correct, na.rm = TRUE) * 100, 4),
    mean_rt      = round(mean(choice.rt[correct == 1],   na.rm = TRUE), 4),
    median_rt    = round(median(choice.rt[correct == 1], na.rm = TRUE), 4),
    sd_rt        = round(sd(choice.rt[correct == 1],     na.rm = TRUE), 4)
  ) %>%
  ungroup() %>%
  mutate(subject = subject_id) %>%
  select(subject, everything()) %>%
  arrange(condition)

# Display summary table
kable(summary_by_cond,
      caption = "Summary of accuracy and RT by condition (all blocks combined)")
```

## 4. Summarize within each block by condition
- Produces the same metrics as above, but broken down by block number.

```{r summarize-by-block}
summary_by_block_and_cond <- dat %>%
  group_by(blocks.thisN, condition) %>%
  summarize(
    n_trials     = n(),
    prop_correct = round(mean(correct, na.rm = TRUE) * 100, 4),
    mean_rt      = round(mean(choice.rt[correct == 1],   na.rm = TRUE), 4),
    median_rt    = round(median(choice.rt[correct == 1], na.rm = TRUE), 4),
    sd_rt        = round(sd(choice.rt[correct == 1],     na.rm = TRUE), 4)
  ) %>%
  ungroup() %>%
  mutate(subject = subject_id) %>%
  select(subject, blocks.thisN, condition, everything()) %>%
  arrange(blocks.thisN, condition)

# Display summary table
kable(summary_by_block_and_cond,
      caption = "Summary of accuracy and RT by block and condition")
```

## 5. Export results to Excel
- Writes both summary tables to an `.xlsx` on your Desktop, named using the subject ID, date, and time.

```{r export-excel}
# Prepare output filename on Desktop
out_fname <- file.path(
  "~/Desktop",
  paste0(
    subject_id, "_", run_date, "_", run_time, "_summary.xlsx"
  )
)

# Write to Excel
write.xlsx(
  list(
    Overall  = summary_by_cond,
    By_Block = summary_by_block_and_cond
  ),
  file = out_fname,
  overwrite = TRUE
)

# Inform where the file is saved
cat("Results exported to:", out_fname, "\n")
```
